Variational Monte Carlo on bosonic systems

### 2018-01-22 project_1/article/article.tex
Add theory on the non-interacting HO system

>>>>

        \subsubsection{Non-interacting harmonic oscillator}
            We start by finding an analytical expression for $a = 0$ and $\beta
            = 1$, i.e., a non-interaction harmonic oscillator given by the
            wavefunction
            \begin{align}
                \bra{\vf{r}}\ket{\Psi_T}
                = \prod_i^N \exp\bigl[
                    -\alpha |\vf{r}_i|^2
                \bigr],
            \end{align}
            where $|\vf{r}_i| = r_i$. As $a = 0$ the interaction term,
            $w(\vf{r}_i, \vf{r}_j)$, vanishes and the Hamiltonian is given by
            (in the spherical case)
            \begin{align}
                H &= \sum_i^N h(\vf{r}_i)
                = \sum_i^N \Bigl(
                    -\frac{\hbar^2}{2m}\nabla_i^2
                    + \half m \omega^2 r_i^2
                \Bigr).
            \end{align}
            Working in spherical coordinates the gradient and the Laplace
            operator (applied to a function) becomes
            \begin{align}
                \nabla f
                &= \vf{e}_r\tpl{r}f
                + \vf{e}_{\theta}\frac{1}{r}\tpl{\theta}f
                + \vf{e}_{\phi}\frac{1}{r\sin(\theta)}\tpl{\phi}f, \\
                \nabla^2 f
                &= \frac{1}{r^2}\tpl{r}
                \bigl[
                    r^2\tpl{r}f
                \bigr]
                + \frac{1}{r^2\sin(\theta)}\tpl{\theta}
                \bigl[
                    \sin(\theta)\tpl{\theta}f
                \bigr] \nonumber\\
                &\qquad
                + \frac{1}{r^2\sin^2(\theta)}\tpl{\phi}^2f,
            \end{align}
            where $\vf{e}_r$ signifies the unit vector in the radial direction
            and we use the shorthand notation
            \begin{align}
                \tpl{r}f \equiv \dpd[]{f}{r}.
            \end{align}
            As the trial wavefunction for $\beta = 1$ is purely radial we get
            the same expression in one, two and three dimensions by setting
            \begin{align}
                r^2 =
                \begin{cases}
                    x & \text{one dimension}, \\
                    x^2 + y^2 & \text{two dimensions}, \\
                    x^2 + y^2 + z^2 & \text{three dimensions}.
                \end{cases}
            \end{align}
            The gradient then becomes
            \begin{align}
                \bra{\vf{r}}\nabla_i\ket{\Psi_T}
                &= \vf{e}_{{r_i}}\tpl{{r_i}}\prod_j^N\exp\bigl[
                    -\alpha r_j^2
                \bigr] \\
                &=
                -2\alpha r_i\prod_j^N\exp\bigl[
                    -\alpha r_j^2
                \bigr]\vf{e}_{{r_i}} \\
                &=
                -2\alpha r_i\bra{\vf{r}}\ket{\Psi_T}\vf{e}_{{r_i}}.
            \end{align}
            The Laplacian is given by
            \begin{align}
                \bra{\vf{r}}\nabla_i^2\ket{\Psi_T}
                &= \frac{1}{r_i^2}\tpl{{r_i}}
                \biggl\{
                    r_i^2\tpl{{r_i}}
                    \prod_j^N\exp\bigl[
                        -\alpha r_j^2
                    \bigr]
                \biggr\} \\
                &=
                \frac{1}{r_i^2}\tpl{{r_i}}
                \biggl\{
                    -2 \alpha r_i^3
                    \prod_j^N\exp\bigl[
                        -\alpha r_j^2
                    \bigr]
                \biggr\} \\
                &=
                \frac{1}{r_i^2}
                \Bigl\{
                    -6 \alpha r_i^2
                    + 4\alpha^2 r_i^4
                \Bigr\}
                \prod_j^N\exp\bigl[
                    -\alpha r_j^2
                \bigr]
                \\
                &=
                \Bigl\{
                    -6 \alpha
                    + 4\alpha^2 r_i^2
                \Bigr\}
                \bra{\vf{r}}\ket{\Psi_T}.
            \end{align}
            This gives the analytical expression for the local energy to be
            \begin{align}
                E_L(\vf{r})
                &= \frac{\bra{\vf{r}}H\ket{\Psi_T}}{\bra{\vf{r}}\ket{\Psi_T}} \\
                &= \sum_i^N \Bigl(
                    4\alpha^2 r_i^2
                    + \half m \omega^2 r_i^2
                \Bigr) - 6\alpha.
            \end{align}
            The \textit{drift force} of the system is given by
            \begin{align}
                \vf{F}_i
                &=
                \frac{2\bra{\vf{r}}\nabla_i\ket{\Psi_T}}
                {\bra{\vf{r}}\ket{\Psi_T}} \\
                &= -4\alpha r_i\vf{e}_{{r_i}}.
            \end{align}
